---
title: "Final Project"
author: "Divy Tripathy"
date: "2024-04-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Impact of Healthcare Availability on Health Outcomes.

I will first explore a complete pooling model. This would imply exchangeablility across entities.

```{r}
library(ggplot2)
library(arm)
library(rstan)
options(mc.cores = parallel::detectCores())
library(TeachingDemos)
library(reshape2)
library(tidyverse)
library(plyr)
library(lme4)
library(rstanarm)
```


```{r}
library(readr)
BLS_Data <- read_csv("~/Developer/VSCode/final-project-3musketeers/data/BLS Data.csv")
BLS_Data <- BLS_Data[-c(1, 2, 53),]
head(BLS_Data)
```

```{r}
CDC_Data <- read_csv("~/Developer/VSCode/final-project-3musketeers/data/CDC Data.csv")
# CDC_2020_Data <- CDC_Data[1:50,]
head(CDC_2020_Data)
```

Below is a varying intercept model with no predictors.

```{r}
vary <- lmer(CDC_Data$RATE ~ (1|CDC_Data$STATE))
summary(vary)
```


```{r}
print(coef(vary)) # mean growth
```

```{r}
print(fixef(vary)) # pooled mean
```

```{r}
print(ranef(vary)) # deviation from pooled mean
```

Unfortunately, not all of the data lends towards hierarchical approaches to to mixed and random effects. In addition, we don't have any "controls". There is more data on the way.


```{r}
library(bayeslm)
pool <- bayeslm(CDC_2020_Data$RATE ~ BLS_Data$Ratio, prior = 'horseshoe', N = 10000, burnin = 2000)
```

```{r}
summary(pool)
```

```{r}
plot(pool$beta)
```
Trace and autocorrelation plots look good.

```{r}
pool_freq <- lm(CDC_2020_Data$RATE ~ BLS_Data$Ratio)
summary(pool_freq)
```


```{r}
plot(BLS_Data$Ratio, CDC_2020_Data$RATE, col="black")
abline(pool_freq, col="blue")
```

Lets look at state level data.

```{r}
utah_final <- read_csv("utah-final.csv")
head(utah_final)
```


```{r}
pool_utah <- bayeslm(utah_final$Deaths ~ utah_final$Ratio, prior = 'horseshoe', N = 10000, burnin = 2000)
summary(pool_utah)
```

```{r}
plot(pool_utah$beta)
```

```{r}
pool_utah <- bayeslm(utah_final$Deaths ~ utah_final$Ratio, prior = 'laplace', N = 10000, burnin = 2000)
summary(pool_utah)
```
```{r}
plot(pool_utah$beta)
```


```{r}
pool_utah_freq <- lm(utah_final$Deaths ~ utah_final$Ratio)
summary(pool_utah_freq)
```


```{r}
plot(utah_final$Ratio, utah_final$Deaths, col = "black")
abline(pool_utah_freq, col = "blue")
```
